A studyforrest extension, an annotation of spoken language in the German dubbed movie ``Forrest Gump'' and its audio-description

### 2020-10-07 paper/main.tex
Cleanup

\author[1, 2]{Christian~O.~Häusler \\ email \href{mailto:c.haeusler@fz-juelich.de}{c.haeusler@fz-juelich.de} }
\author[1, 2]{Michael~Hanke \\ email \href{mailto:m.hanke@fz-juelich.de}{m.hanke@fz-juelich.de} }

@@

\todo[inline]{wie email und ORCID (du: https://orcid.org/0000-0001-6398-6370; ich: https://orcid.org/0000-0002-0936-317X) unterbringen?}

@@

\begin{abstract}
% up to 300 words intro
Here we present an annotation of speech spoken in the audio-visual movie
``Forrest Gump'' and its audio-description (i.e. the audio-only variant of the
movie) as an addition to a large public functional brain imaging dataset
(\href{www.studyforrest.org}{studyforrest.org}).
% anno content 1
The annotation provides information about the exact timing of each of the more
than 2500 spoken sentences, 16000 words (including 202 non-speech
vocalizations), 66000 phonemes, and their corresponding speaker.
% anno content 2
Additionally for every word, we provide lemmatization, a simple
part-of-speech-tagging (15 grammatical categories), a detailed part-of-speech
tagging (43 grammatical categories), syntactic dependencies, and a semantic
analysis based on word embedding which represents each word in a
300-dimensional semantic space.
% validation
To validate the dataset's quality, we build a model of hemodynamic brain
activity based on information drawn from the annotation.
% why bother
Results suggest that the annotation's content and quality enable independent
researchers to create sophisticated models of brain activity correlating with a
variety of linguistic aspects under conditions of real-life complexity.
\end{abstract} \todo{validation is now (carefully) mentioned; 150 of 300 words
used for abstract}

@@

\subsection*{Dataset content}
% \href{https://bids.neuroimaging.io/}{bids.neuroimaging.io} TextGrid format
The annotation comes in two different versions. First, as a text-based TextGrid
file (\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.TextGrid}) to be
conveniently edited using the software Praat \citep{boersma2019praat}.
% TSV format
Second, as a text-based, tab-separated-value (TSV) formatted table
(\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.tsv}) in accordance with
the brain imaging data structure (\href{https://bids.neuroimaging.io/}{BIDS})
\citep{gorgolewski2016bids}.
% source code for descriptive statistics
The source code for all descriptive statistics included in this paper is
available in \texttt{code/descriptive-statistics.py} (Python script).

\section*{Dataset validation}
\todo[inline]{``Information about any validation carried out and/or any
    limitations of the datasets, including any allowances made for controlling
bias or unwanted sources of variability. ~170 words'' 170 words? lol}
% intro
In order to assess the annotation's quality, we investigated if contrasting
speech-related events to events without speech lead to increased activation in
areas known to be involved in language processing \citep{hickok2007cortical}.
% nouns and proper nouns resp. > coordinate junctions
Moreover, we tested if two similar linguistic concepts (proper nouns and nouns)
providing high semantic information contrasted with a concept providing low
semantic information (coordinate conjunctions) lead to increased activation in
congruent brain areas.

@@

% Discussion words > no-speech
For the contrast words > no-speech, results show increased hemodynamic activity
in a bilateral cortical network including temporal, parietal and frontal regions
related to processing spoken language \citep{friederici2011brain,
hickok2007cortical,price2012twentyyears}.
% similarities to previous naturalistic speech paradigms
These clusters resemble results of previous studies that implemented an ISC
approach to analyze fMRI data of naturalistic auditory stimuli
\citep{honey2012not, lerner2011topographic, silbert2014coupled}.
% differences
We do not find significantly increased activations in midline areas (like the
posterior cingulate cortex and precuneus or anterior cingulate cortex and medial
frontal cortex) which showed synchronized activity across subjects in previous
studies.
% Wilson (2008): ISC + GLM
In this regard, our results are similar to \citep{wilson2008beyond} who
implemented both an ISC and a GLM analysis. In this study, the ISC analysis
showed synchronized activity in midline areas but the GLM analysis contrasting
blocks of listening to narratives to blocks of a resting condition showed
significantly decreased activity in these areas.
\todo[inline]{our reverse contrast (no-speech > words): clusters in posterior \&
anterior cingulate (and subcallosal c. \& insulae bilaterally); I think opening
the pandora's box of the reverse contrasts is not worth it here}

% (proper) nouns > con
The two contrasts that contrasted nouns and proper nouns respectively to
coordinate junctions yielded increased activation partially located in early
sensory regions (Heschl's Gyrus; \citep{saenz2014tonotopic}) and most
prominently adjacent regions bilaterally (planum temporale; superior temporal
gyrus; \citep{arsenault2015distributed, mesgarani2014phonetic}).
% why these contrasts nous & proper nouns
We chose nouns and proper nouns for these two contrasts because they represent
linguistically similar concepts but are yet uncorrelated in the German language
and stimulus (cf. Figure \ref{fig:reg-corr}).
% coordinate conjunctions
We contrasted nouns and proper nouns respectively to coordinate conjunctions
because nouns and proper nouns are linguistically different to coordinate
conjunctions as well as uncorrelated.
% “interpretation”
\todo[inline]{well what's coming up next is adventurous} Despite the fact that
nouns and proper nouns are uncorrelated, both contrasts lead to largely
spatially congruent clusters. Results suggest that models based on our
annotation of similar linguistic concepts correlate with hemodynamic activity in
spatially similar areas.
% not driven by coordinate conjunction regressors
We confirmed the validity of these interpretation by testing if the spatial
congruency could be attributed to a negative correlation of coordinate
conjunctions with the modeled time series which turned out not to be the case.
% why it's cool
In summary, results of our exploratory analyses suggest that the annotation of
speech meets the quality requirements to be used as the basis for model-based
analyses that investigate language perception under more ecologically valid
conditions.\todo{might be a little humble}

@@

\section*{Conclusions}
\todo[inline]{give an accurate assessment what this can be used for and what is
    problematic or impossible.  ATM this is the only statement in this regard:
    'facilitates modeling of hemodynamic brain responses correlating with
    speech-related events' and the only indication of purpose is 'additional
confound measures'. I thin this is incongruent with the amount of work and the
level of detail of this annotation; CH: let's discuss that personally}
% what the paper is about
As an extension of the studyforrest dataset, we present an annotation of spoken
language in the two hours lasting ``research cut'' of the audio-visual movie
``Forrest Gump'' and its temporally aligned audio-description.
% what we did
We annotated the onset and offset of sentences, words and phonemes, and the
corresponding speaker's identity.
% what for
An additional tagging of each word's linguistic features enables independent
working groups to create sophisticated models of hemodynamic activity
correlating with linguistics aspects like phonetics, grammar, syntax, and
semantics occurring in two naturalistic stimuli.
% results
Results of our analysis that we ran to assess the dataset's quality shows that
the annotation's content can indeed be used to isolate speech-related networks
in the human brain under life-like conditions.

>>>>

\author[1, 2]{Christian~O.~Häusler}
\author[1, 2]{Michael~Hanke}

@@

@@

\begin{abstract}
% up to 300 words intro
Here we present an annotation of speech spoken in the audio-visual movie
``Forrest Gump'' and its audio-description (i.e. the audio-only variant of the
movie) as an addition to a large public functional brain imaging dataset
(\href{www.studyforrest.org}{studyforrest.org}).
% anno content 1
The annotation provides information about the exact timing of each of the more
than 2500 spoken sentences, 16000 words (including 202 non-speech
vocalizations), 66000 phonemes, and their corresponding speaker.
% anno content 2
Additionally for every word, we provide lemmatization, a simple
part-of-speech-tagging (15 grammatical categories), a detailed part-of-speech
tagging (43 grammatical categories), syntactic dependencies, and a semantic
analysis based on word embedding which represents each word in a
300-dimensional semantic space.
% validation
To validate the dataset's quality, we build a model of hemodynamic brain
activity based on information drawn from the annotation.
% why bother
Results suggest that the annotation's content and quality enable independent
researchers to create models of brain activity correlating with a
variety of linguistic aspects under conditions of real-life complexity.
\end{abstract}

\section*{Contacts}
\begin{description}
  \item{Christian~O.~Häusler} \\
    email: \href{mailto:c.haeusler@fz-juelich.de}{c.haeusler@fz-juelich.de} \\
    orcid: https://orcid.org/0000-0002-0936-317X
  \item{Michael~Hanke}\\
    email: \href{mailto:michael.hanke@gmail.com}{michael.hanke@gmail.com} \\
    orcid: https://orcid.org/0000-0001-6398-6370 \\
\end{description}

@@

\subsection*{Dataset content}
% \href{https://bids.neuroimaging.io/}{bids.neuroimaging.io} TextGrid format
The annotation comes in two different versions. First, as a text-based TextGrid
file (\texttt{annotation/ fg\_rscut\_ad\_ger\_speech\_tagged.TextGrid}) to be
conveniently edited using the software Praat \citep{boersma2019praat}.
% TSV format
Second, as a text-based, tab-separated-value (TSV) formatted table
(\texttt{annotation/fg\_rscut\_ ad\_ger\_speech\_tagged.tsv}) in accordance with
the brain imaging data structure (\href{https://bids.neuroimaging.io/}{BIDS})
\citep{gorgolewski2016bids}.
% source code for descriptive statistics
The source code for all descriptive statistics included in this paper is
available in \texttt{code/descriptive-statistics.py} (Python script).

\section*{Dataset validation}
% intro
In order to assess the annotation's quality, we investigated if contrasting
speech-related events to events without speech lead to increased activation in
areas known to be involved in language processing \citep{hickok2007cortical}.
% nouns and proper nouns resp. > coordinate junctions
Moreover, we tested if two similar linguistic concepts (proper nouns and nouns)
providing high semantic information contrasted with a concept providing low
semantic information (coordinate conjunctions) lead to increased activation in
congruent brain areas.

@@

% Discussion words > no-speech
For the contrast words > no-speech, results show increased hemodynamic activity
in a bilateral cortical network including temporal, parietal and frontal regions
related to processing spoken language \citep{friederici2011brain,
hickok2007cortical,price2012twentyyears}.
% similarities to previous naturalistic speech paradigms
These clusters resemble results of previous studies that implemented an ISC
approach to analyze fMRI data of naturalistic auditory stimuli
\citep{honey2012not, lerner2011topographic, silbert2014coupled}.
% differences
We do not find significantly increased activations in midline areas (like the
posterior cingulate cortex and precuneus or anterior cingulate cortex and medial
frontal cortex) which showed synchronized activity across subjects in previous
studies.
% Wilson (2008): ISC + GLM
In this regard, our results are similar to \citep{wilson2008beyond} who
implemented both an ISC and a GLM analysis. In this study, the ISC analysis
showed synchronized activity in midline areas but the GLM analysis contrasting
blocks of listening to narratives to blocks of a resting condition showed
significantly decreased activity in these areas.

% (proper) nouns > con
The two contrasts that contrasted nouns and proper nouns respectively to
coordinate junctions yielded increased activation partially located in early
sensory regions (Heschl's Gyrus; \citep{saenz2014tonotopic}) and most
prominently adjacent regions bilaterally (planum temporale; superior temporal
gyrus; \citep{arsenault2015distributed, mesgarani2014phonetic}).
% why these contrasts nous & proper nouns
We chose nouns and proper nouns for these two contrasts because they represent
linguistically similar concepts but are yet uncorrelated in the German language
and stimulus (cf. Figure \ref{fig:reg-corr}).
% coordinate conjunctions
We contrasted nouns and proper nouns respectively to coordinate conjunctions
because nouns and proper nouns are linguistically different to coordinate
conjunctions as well as uncorrelated.
% “interpretation”
Despite the fact that
nouns and proper nouns are uncorrelated, both contrasts lead to largely
spatially congruent clusters. Results suggest that models based on our
annotation of similar linguistic concepts correlate with hemodynamic activity in
spatially similar areas.
% not driven by coordinate conjunction regressors
We confirmed the validity of these interpretation by testing if the spatial
congruency could be attributed to a negative correlation of coordinate
conjunctions with the modeled time series which turned out not to be the case.
% why it's cool
In summary, results of our exploratory analyses suggest that the annotation of
speech meets the quality requirements to be used as the basis for model-based
analyses that investigate language perception under more ecologically valid
conditions.

@@

\section*{Conclusions}
% what the paper is about
As an extension of the studyforrest dataset, we present an annotation of spoken
language in the two hours lasting ``research cut'' of the audio-visual movie
``Forrest Gump'' and its temporally aligned audio-description.
% what we did
We annotated the onset and offset of sentences, words and phonemes, and the
corresponding speaker's identity.
% what for
An additional tagging of each word's linguistic features enables independent
working groups to create models of hemodynamic activity
correlating with linguistics aspects like phonetics, grammar, syntax, and
semantics occurring in two naturalistic stimuli.
% results
Results of our analysis that we ran to assess the dataset's quality shows that
the annotation's content can indeed be used to isolate speech-related networks
in the human brain under life-like conditions.
